### 2011-03-26 Review/answers.tex
Working with corrections.

\textbf{Question:}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estima
tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.

\textbf{Answer:} Thank you for the suggestion. While we agree that there is redundant information
about initialization and this was taken into the account by reducing repetitive text:
\begin{quotation}
\end{quotation}

 However, the
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
keep equation 16 as the most informative for the LBM community and to reduce {\color{red} equation
14 and 6}.
\bibliographystyle{unsrtnat}
\bibliography{answers}

>>>>

\textbf{Question:}
\begin{quotation}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estima
tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.
\end{quotation}

\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
about initialization and this was taken into the account by reducing repetitive text and
introducing step-by-step explations, see Section about initialization. We also added the
recommendations for controlling parameters in the same section.  The
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
keep suggestions for controlling and initialization of simulations as informative guides for the
LBM community.

\textbf{Question:}
\begin{quotation}
 Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
obtained
under the assumption of an axisymmetric Poiseuille ﬂow. However, among the ”Key words”, one can
read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
that are done in this work. The Authors should discuss this choice of estimating the pressure
gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
in Equation 5).
\end{quotation}

\textbf{Answer:} While the Hele-Shaw flow implies the flow with the infinitisemal distance between
plates, in our situation the size of the bubble can be compared with the channel height ($5$ times
larger) and that's why the assumption of the infinitisemal distance between plates is too strict.
Moreover, after recalculating results with the velocity of the interface the assumption of the
Poiseulle flow is valid and shows good accuracy.

\textbf{Question:}
\begin{quotation}
The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
journal of applied science. The Authors would improve signiﬁcantly the manuscript by deﬁning more
rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
A suggestion is to emphasize within an array the links between physical parameters and the model
eq
parameters. The variables i, fieq , gi should be deﬁned and the Equations for f and g in Equation
3 should be described (which equation corresponds to continuity equation...). Equation 2 could
perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
physical variable f of Equation 3. The notation i = 1 ÷ 8 is unusual, it would be clearer to use a
classical notation such as 1 ≤ i ≤ 8. The sound speed c2 should be deﬁned just after its
introduction in line 107.
\end{quotation}

\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
parameters with the lattice Botlzmann parameters. As well more explanations are put towards the
interface surface tension description and inclusion it to the model. The reference to the
continuity and the Navier-Stokes equations are added. The equation (2) is not rewritten in terms of
generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
governing equations. The notations are corrected.

\textbf{Question:}
\begin{quotation}
The name of the Section ”Lattice Boltzmann benchmark” lets the reader think that this Section
would describe the numerical procedure (assumptions, choice of the geometry and the grid, boundary
conditions, control parameters, initialization of the scheme). This is not the case because this
Section
develops the challenges and the difficulties of using the Lattice Boltzmann method to bubble ﬂows.
This Section should be renamed or reorganized.
\end{quotation}

\textbf{Answer:} We renamed the section to ``Approach to benchmark''.

\textbf{Question:}
\begin{quotation}
The Authors should make a distinction between variables in lattice units and dimensional or non
dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y − 2$ where $N_y$ is
the
number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
is in
contradiction with the use of $H_{eff}$ in Equation 5. The Authors should deﬁne each variable
without
any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
improvements
would ﬁt well in the Subsection named ”The nondimensionalization and initialization procedure”
which does not in fact introduce much nondimensionalisation procedure.
\end{quotation}

\textbf{Answer:} We added a decription paragraph in the Subsection ``The nondimensionalization and
initialization procedure'' about the approach to units. In fact, all equations are the same in the
lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
preferred units as soon as capillary number is matched in the LBM and the physical world.

\textbf{Question:}
\begin{quotation}
Some notations are unclear. For instance, the ﬁlm thickness is denoted as h∞ in Equation 7 whereas
it is denoted δ in most of the paper. The notations w or δsim are also found before line 381 or in
Table 3 respectively. The thickness of the interface is denoted ξ in Figure 1 but the same variable
is
denoted 5ξ in Table 1. The speed of the bubble is denoted Ububble in some places and Ub in others.
The variable Uin Equation 5 does not seem to be deﬁned.
\end{quotation}

\texbf{Answer:} We thoroughly corrected all the inconsistencies.

\bibliographystyle{unsrtnat}
\bibliography{answers}
